---
title: "Analysis - GY"
author: "Gavin Yoesting"
date: "5/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Analysis of Code Languages
The purpose of this project is to compare and analyze data using three different coding languages: R, Mathematica, and Python. The data used in this project was taken from a general chemistry diffraction lab held at Oklahoma Baptist University. Four data sets from several different students were inputted to complete this project. The data used analyzes fringes, the diameter of hair, and the overall average diameter of hair. 

## R
R is a free programming software used for statistical computing and graphics. It is predominantly used among statisticians and data miners for data analysis. In this project, the data was inputted, plotted, and analyzed by Alexis Gilliam until pushed into Git.

# Inputting the Data
To start using the program data must be present. The data being used is currently in another R Markdown document but must be added to this one as well in a readable format. There are many different approaches to this. For our purposes this data is presented as a data frame.

This first set of data show the individual fringe, diameter, wavelength, and distance.
```{r}
diffractiondata <- read.table(header=TRUE, text="
Fringe   Diameter(um) Wavelength(nm)   Distance(cm)   
1        88.64        650              2.2
1        97.5         650              2.0
1        43.33        650              1.5
1        58.56        640              1.0
2        86.67        650              4.5
2        86.67        650              4.5
2        52.00        650              2.5
2        58.56        640              2.0
3        82.39        650              7.1
3        90.0         650              6.5
3        51.3         650              3.8
3        79.85        640              2.2
4        84.78        650              9.2
4        91.76        650              8.5
4        54.16        650              4.8
4        83.65        640              2.8
average  79.32        648              N/A")    
```

This second data frame measures the average diameter of each hair color. 
```{r}
averagediameter <- read.table(header=TRUE, text="
Group     Average_Diameter(um)   Color
1           196.78               brown
2           96.4                 blonde
3           161.76               brown
4           85.62                brown
5           30.58                red
6           122.8                red")
```

Data vectors for line charts
```{r}
Firstfringediameterum <- c(88.64, 97.5, 43.33, 58.56)
Firstfringediameterum
Firstfringedistancecm <- c(2.2, 2.0, 1.5, 1.0)
Firstfringedistancecm
Secondfringediameterum <- c(86.67, 86.67, 52.00, 58.56)
Secondfringediameterum
Secondfringedistancecm <- c(4.5, 4.5, 2.5, 2.0)
Secondfringedistancecm
Thirdfringediameterum <- c(82.39, 90.0, 51.3, 79.85)
Thirdfringediameterum
Thirdfringedistancecm <- c(7.1, 6.5, 3.8, 2.2)
Thirdfringedistancecm
Fourthfringediamterum <- c(84.78, 91.76, 54.16, 83.65)
Fourthfringediamterum
Fourthfringedistancecm <- c(9.2, 8.5, 4.8, 2.8)
Fourthfringedistancecm
averagediameters <- c(196.78, 96.4, 161.76, 85.62, 30.58, 122.8)
averagediameters
```

# Plotting Data 
In the start of analysis, the data of the first data frame is analyzed in a scatter plot. The color of each of the fringes is changed to allow easy comparison.

```{r}
library(ggplot2)
ggplot(diffractiondata, aes(x = Distance.cm., y = Diameter.um.)) + 
  geom_point(aes(color = factor(Fringe)))
```

This same comparison can be drawn for the second data frame which brings into question the actually hair colors.

```{r}
library(ggplot2)
ggplot(averagediameter, aes(x = Group, y = Average_Diameter.um.)) +
  geom_point(aes(color = factor(Color)))
```

This makes it easy to compare which groups worked with which hair color and what their results are compared to each other. 

Data can also be observed in line graphs, but only specific data
```{r}
a <- averagediameters
plot(a, type = "o", col = "blue", xlab = "Group Number", ylab = "Diameter Measured in um", main = "Average Diameters of Hair by Group")
```

# Data Tables

Data can be organized in tables
```{r}
read.csv(file = "DiffractionDataTable1.csv", header = TRUE)
```

# Analysis of of Data
Setting up the groups of data
At this point we have already set up each of the data strings into a single command. We can take these vectors and create combined groups of them.
```{r}
fringediameterall <- data.frame(cbind(Firstfringediameterum, Secondfringediameterum, Thirdfringediameterum, Fourthfringediamterum))
fringediameterall
summary(fringediameterall) #This shows the mean, median, minimum, and maximums through each of these data sets.
```
This can be repeated with the average diameter, however we don't need to put it in a combined group.
```{r}
summary(averagediameter)
```
We can take ANOVAs for each of these data sets as well.
```{r}
 fringediameter<- stack(fringediameterall)
fringediameter
Anova_fringediameter <- aov(values ~ ind, data = fringediameter)
summary(Anova_fringediameter)
```

## Mathematica
Mathematica is a symbolic mathematical computation program, sometimes called a computer algebra program, used in many scientific, engineering, mathematical, and computing fields. This file will contain the URL for the completed mathematica coding. The data was inputted, plotted, and analyzed by Miranda Plaster until pushed to Git. This URL helps allow this file to be easily viewed in its entirety.URL: https://www.wolframcloud.com/obj/eee32802-7a0e-4018-bc3e-50b66f126d4e

## Python
Python is a multi-paradigm programming language. Many software engineers and researchers use this program because of its code readability and significant white space. In this project, the data was inputted, plotted, and analyzed by Nick Sitter until pushed to Git.

# Inputting the Data
This file was written to code different processes such as calculating averages, creating boxplots, and performing an ANOVA. Notable tables can be seen below.

Table 1. Combined data of all trials completed

diffractiondata = read.table(header=TRUE, text="
Fringe   Diameter(um) Wavelength(nm)   Distance(cm)   
1        88.64        650              2.2
1        97.5         650              2.0
1        43.33        650              1.5
1        58.56        640              1.0
2        86.67        650              4.5
2        86.67        650              4.5
2        52.00        650              2.5
2        58.56        640              2.0
3        82.39        650              7.1
3        90.0         650              6.5
3        51.3         650              3.8
3        79.85        640              2.2
4        84.78        650              9.2
4        91.76        650              8.5
4        54.16        650              4.8
4        83.65        640              2.8
average  79.32        648              N/A") 

Table 2. Average Diameter by Hair Color
 
Group     Average Diameter(um)   Color
1           196.78               brown
2           96.4                 blonde
3           161.76               brown
4           85.62                brown
5           30.58                red
6           122.8                red")

Code Block 1 - Averages Diameters of Fringe 1
```{python}
def avg(numList):
    avg = sum(numList) / len(numList)
    return avg

Fringe1 = [88, 97.5, 43.33, 58.56]
print(avg(Fringe1))
```

Code Block 2 - Averages Diameters of Fringe 2
```{python}
def avg(numList):
    avg = sum(numList) / len(numList)
    return avg

Fringe2 = [86.67, 86.67,52, 58.56]
print(avg(Fringe2))
```

Code Block 3 - Averages Diameters of Fringe 3
```{python}
def avg(numList):
    avg = sum(numList) / len(numList)
    return avg

Fringe3 = [82.39, 90, 51.3, 79.85]
print(avg(Fringe3))
```

Code Block 4 - Averages Diameters of Fringe 4
```{python}
def avg(numList):
    avg = sum(numList) / len(numList)
    return avg

Fringe4 = [84.78, 91.76, 54.16, 83.65]
print(avg(Fringe4))
```

# Plotting the Data
Code Block 5 - Creates a png File of a Boxplot
```{python}
import matplotlib as mpl 
mpl.use('agg')
import matplotlib.pyplot as plt 
Fringe1 = (88, 97.5, 43.33, 58.56)
Fringe2 = (86.67, 86.67,52, 58.56)
Fringe3 = (82.39, 90, 51.3, 79.85)
Fringe4 = (84.78, 91.76, 54.16, 83.65)
data_to_plot = [Fringe1, Fringe2, Fringe3, Fringe4]
fig = plt.figure(1, figsize=(9, 6))
ax = fig.add_subplot(111)
boxplot = ax.boxplot(data_to_plot)
fig.savefig('fig1.png', bbox_inches='tight')
```

# Data Tables
Code Block 6 - Gives ANOVA Table Between The Different Fringes
```{python}
import statsmodels.api as sm 
from statsmodels.formula.api import ols
import pandas as pd
datafile = "DiffractionDataTable1.csv"
data = pd.read_csv(datafile)
mod = ols('Fringe ~ Diameter', data=data).fit()
aov_table = sm.stats.anova_lm(mod, typ=2)
print(aov_table)
```

## Analysis of R
The analysis of data from R was very useful in the overall project. The coding seems to be a very long and time consuming process compared to the other programs, but provides in depth applications such as the mean, median, mode, maximum, and minimum for each specific fringe. The process of inputting diffraction data and data vectors that take over ten lines is strenuous compared to the simple code blocks Mathematica and Python used. The R code blocks also took the majority of the overall chunk time to run. Although time into inputting and running data detract from its user friendliness, its output of data tables and figures were more informational. Three graphs were plotted: the diffraction data in a scatter plot with color coded fringes, the average diameter in a scatter plot with color coded fringes, and the average diameter of hair by group in a line graph. Data is more suited for statistical support and data analysis by using R. Factors such as the mean, median, mode, maximum, and minimum were also avaliable while Mathematica and Python only provided specific factors such as the average when run. 

Overall, R required more code blocks and commands, but gave adequate and professional styled data that was the most useful compared to Python and Mathematica. 

## Analysis of Mathematica
The analysis of data from Mathematica was useful, but not as in depth as the R program for data analysis. It contained a list plot and a line plot as visuals to see the data. The list plot is identical to the scatter plot provided in R, but the data is inputted in a more user friendly way with less code lines used. It included listing the data with the distance as the x-coordinate and the diameter as the y-coordinate with sixteen data pairs. The line graph used the same user friendly process, but gave an unusual line in the graph. Mathematica also provided two data tables: diffraction data listed by fringe table and a find anomolies table. These can be useful for a quick analysis, but are not as useful to the R graphs. The t test seemed to function irregularly due to the data in Mathematica and did not provide adequate date. Compared to Python. it has a similar simple language but is more useful for the usage of line plots. Python gives more data regarding the minimum, maximum, and average. 

Overall, Mathematic has a simple language that is easy to learn, but provides quick and basic data compared to R. The program's graphing capability for a visualization of data is more useful for line graphs than Python which provides only a graph with the maximum, medium, and minimum. 

## Analysis of Python
The analysis of data from Python gave specific factors such as the average diameter of each fringe, a graph that outlines the average, maximum, and minimum, and a data table. It has a simple language like Mathematica, but focuses more on the average, maximum, and minimum in fig1.png. This program also can be used for quick data, but may lack some information that can be provided in R.

Overall, Python is not as useful as R and is more object-oriented. The program does have some benefits such as the minimum, maximum, and average graph compared to Mathematica.

## Conclusion
This project helped me understand each program more in depth with the analysis of data. Each program has supportive functions for analyzing data, but R is the most informative. The use of three programs also helped me learn more about coding, specifically about downloading packages that were needed. Overall, this project helped me understand the importance of programming software.