---
title: "Analysis_Of_Data_Languages"
author: "Kayla Tumbaga"
date: "May 3, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Code
This first set of data show the individual fringe, diameter, wavelength, and distance.
```{r}
diffractiondata <- read.table(header=TRUE, text="
Fringe   Diameter(um) Wavelength(nm)   Distance(cm)   
1        88.64        650              2.2
1        975          650              2.0
1        43.33        650              1.5
1        58.56        640              1.0
2        86.67        650              4.5
2        86.67        650              4.5
2        52.00        650              2.5
2        58.56        640              2.0
3        82.39        650              7.1
3        90.0         650              6.5
3        51.3         650              3.8
3        79.85        640              2.2
4        84.78        650              9.2
4        91.76        650              8.5
4        54.16        650              4.8
4        83.65        640              2.8
average  79.32        648              N/A")    
```

This second data frame measures the average diameter of each hair color. 
```{r}
averagediameter <- read.table(header=TRUE, text="
Group     Average_Diameter(um)   Color
1           196.78               brown
2           96.4                 blonde
3           161.76               brown
4           85.62                brown
5           30.58                red
6           122.8                red")
```

Data vectors for line charts
```{r}
Firstfringediameterum <- c(88.64, 975, 43.33, 58.56)
Firstfringediameterum
Firstfringedistancecm <- c(2.2, 2.0, 1.5, 1.0)
Firstfringedistancecm
Secondfringediameterum <- c(86.67, 86.67, 52.00, 58.56)
Secondfringediameterum
Secondfringedistancecm <- c(4.5, 4.5, 2.5, 2.0)
Secondfringedistancecm
Thirdfringediameterum <- c(82.39, 90.0, 51.3, 79.85)
Thirdfringediameterum
Thirdfringedistancecm <- c(7.1, 6.5, 3.8, 2.2)
Thirdfringedistancecm
Fourthfringediamterum <- c(84.78, 91.76, 54.16, 83.65)
Fourthfringediamterum
Fourthfringedistancecm <- c(9.2, 8.5, 4.8, 2.8)
Fourthfringedistancecm
averagediameters <- c(196.78, 96.4, 161.76, 85.62, 30.58, 122.8)
averagediameters
```


# Plotting Data 

In the start of analysis, the data of the first data frame is analyzed in a scatter plot. The color of each of the fringes is changed to allow easy comparison.

```{r}
library(ggplot2)
ggplot(diffractiondata, aes(x = Distance.cm., y = Diameter.um.)) + 
  geom_point(aes(color = factor(Fringe)))
```

This same comparison can be drawn for the second data frame which brings into question the actually hair colors.

```{r}
library(ggplot2)
ggplot(averagediameter, aes(x = Group, y = Average_Diameter.um.)) +
  geom_point(aes(color = factor(Color)))
```

This makes it easy to compare which groups worked with which hair color and what their results are compared to each other. 

Data can also be observed in line graphs, but only specific data
```{r}
a <- averagediameters
plot(a, type = "o", col = "blue", xlab = "Group Number", ylab = "Diameter Measured in um", main = "Average Diameters of Hair by Group")
```

# Data Tables

Data can be organized in tables
```{r}
read.csv(file = "fringesanddiameters.csv", header = TRUE)
```

# Analysis of of Data

Setting up the groups of data
At this point we have already set up each of the data strings into a single command. We can take these vectors and create combined groups of them.
```{r}
fringediameterall <- data.frame(cbind(Firstfringediameterum, Secondfringediameterum, Thirdfringediameterum, Fourthfringediamterum))
fringediameterall
summary(fringediameterall) #This shows the mean, median, minimum, and maximums through each of these data sets.
```
This can be repeated with the average diameter, however we don't need to put it in a combined group.
```{r}
summary(averagediameter)
```
We can take ANOVAs for each of these data sets as well.
```{r}
 fringediameter<- stack(fringediameterall)
fringediameter
Anova_fringediameter <- aov(values ~ ind, data = fringediameter)
summary(Anova_fringediameter)
```


##Python Code
Table 1. Combined data of all trials completed

diffractiondata = read.table(header=TRUE, text="
Fringe   Diameter(um) Wavelength(nm)   Distance(cm)   
1        88.64        650              2.2
1        975          650              2.0
1        43.33        650              1.5
1        58.56        640              1.0
2        86.67        650              4.5
2        86.67        650              4.5
2        52.00        650              2.5
2        58.56        640              2.0
3        82.39        650              7.1
3        90.0         650              6.5
3        51.3         650              3.8
3        79.85        640              2.2
4        84.78        650              9.2
4        91.76        650              8.5
4        54.16        650              4.8
4        83.65        640              2.8
average  79.32        648              N/A") 

```{python}
def avg(numList):
    avg = sum(numList) / len(numList)
    return avg

DataSetA = [88, 975, 43.33, 58.56]
print(avg(DataSetA))
```

```{python}
def avg(numList):
    avg = sum(numList) / len(numList)
    return avg

DataSetB = [86.67, 86.67,52, 58.56]
print(avg(DataSetB))
```

```{python}
def avg(numList):
    avg = sum(numList) / len(numList)
    return avg

DataSetC = [82.39, 90, 51.3, 79.85]
print(avg(DataSetC))
```

```{python}
def avg(numList):
    avg = sum(numList) / len(numList)
    return avg

DataSetD = [84.78, 91.76, 54.16, 83.65]
print(avg(DataSetD))
```
d = read.table(header=TRUE, text="
A	    B	    C	    D
88	  86.67	82.39	84.78
975	  86.67	90.0	91.76
43.33	52	  51.3	54.16
58.56	58.56	79.85	83.65

```{python}
d.boxplot(column=['A', 'B', 'C', 'D'], grid=False)
```

##Mathematica Coding
This file will contain the URL for the completed mathematica coding. 
This is in order to allow this file to be easily viewed in its entirety.
URL: https://www.wolframcloud.com/obj/eee32802-7a0e-4018-bc3e-50b66f126d4e

##Analysis Between R Code and Mathematica
Using R coding rather than Mathematica the information gathered
experimentally can by kept in a tidy data format with a clear way
of identifying the data sets that are being analyzed. In
mathematica the table which contains the data set is unclear as to
what data is being analyzed. Mathematica however allows one to mix
symbollic and numberical calculations in their data sets and by
having fewer abbreviated commands the code is more readable.
However, R allows for interaction with a variety of different
programs as there is a variety of different programs that can be
downloaded from the open source network of R that allows for
further data analysis such as ggplot2. R tends to run on an
element by element basis whereas mathematica is more generalized
in terms of multi argument functions.Due to the simplified code 
language, in mathematica models can be generated quicker than they
can be with R as each element does not have to be defined for the
code to function. 

##Analysis Between R Code and Python
Both R and Python are open source programming languages which 
allows for a variety of different programs to interact when 
analyzing a data set as packages can be downloaded to analyze data
in a variety of different ways. 
##Analysis Between Python and Mathematica

##Conclusion 