---
title: "R code"
author: "Alexis Gilliam"
date: "May 2, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Inputting the Data
To start using the program data must be present. The data being used is currently in another R Markdown document but must be added to this one as well in a readable format. There are many different approaches to this.For our purposes this data is presented as a data frame.

This first set of data show the individual fringe, diameter, wavelength, and distance.
```{r}
diffractiondata <- read.table(header=TRUE, text="
Fringe   Diameter(um) Wavelength(nm)   Distance(cm)   
1        88.64        650              2.2
1        975          650              2.0
1        43.33        650              1.5
1        58.56        640              1.0
2        86.67        650              4.5
2        86.67        650              4.5
2        52.00        650              2.5
2        58.56        640              2.0
3        82.39        650              7.1
3        90.0         650              6.5
3        51.3         650              3.8
3        79.85        640              2.2
4        84.78        650              9.2
4        91.76        650              8.5
4        54.16        650              4.8
4        83.65        640              2.8
average  79.32        648              N/A")    
```

This second data frame measures the average diameter of each hair color. 
```{r}
averagediameter <- read.table(header=TRUE, text="
Group     Average_Diameter(um)   Color
1           196.78               brown
2           96.4                 blonde
3           161.76               brown
4           85.62                brown
5           30.58                red
6           122.8                red")
```

Data vectors for line charts
```{r}
Firstfringediameterum <- c(88.64, 975, 43.33, 58.56)
Firstfringediameterum
Firstfringedistancecm <- c(2.2, 2.0, 1.5, 1.0)
Firstfringedistancecm
Secondfringediameterum <- c(86.67, 86.67, 52.00, 58.56)
Secondfringediameterum
Secondfringedistancecm <- c(4.5, 4.5, 2.5, 2.0)
Secondfringedistancecm
Thirdfringediameterum <- c(82.39, 90.0, 51.3, 79.85)
Thirdfringediameterum
Thirdfringedistancecm <- c(7.1, 6.5, 3.8, 2.2)
Thirdfringedistancecm
Fourthfringediamterum <- c(84.78, 91.76, 54.16, 83.65)
Fourthfringediamterum
Fourthfringedistancecm <- c(9.2, 8.5, 4.8, 2.8)
Fourthfringedistancecm
averagediameters <- c(196.78, 96.4, 161.76, 85.62, 30.58, 122.8)
averagediameters
```


# Plotting Data 

In the start of analysis, the data of the first data frame is analyzed in a scatter plot. The color of each of the fringes is changed to allow easy comparison.

```{r}
library(ggplot2)
ggplot(diffractiondata, aes(x = Distance.cm., y = Diameter.um.)) + 
  geom_point(aes(color = factor(Fringe)))
```

This same comparison can be drawn for the second data frame which brings into question the actually hair colors.

```{r}
library(ggplot2)
ggplot(averagediameter, aes(x = Group, y = Average_Diameter.um.)) +
  geom_point(aes(color = factor(Color)))
```

This makes it easy to compare which groups worked with which hair color and what their results are compared to each other. 

Data can also be observed in line graphs, but only specific data
```{r}
a <- averagediameters
plot(a, type = "o", col = "blue", xlab = "Group Number", ylab = "Diameter Measured in um", main = "Average Diameters of Hair by Group")
```

# Data Tables

Data can be organized in tables
```{r}
read.csv(file = "C:/Users/alexi/Downloads/Biochemlab/Comp Project 2/fringesanddiameters.csv", header = TRUE)
```

# Analysis of of Data

Setting up the groups of data
At this point we have already set up each of the data strings into a single command. We can take these vectors and create combined groups of them.
```{r}
fringediameterall <- data.frame(cbind(Firstfringediameterum, Secondfringediameterum, Thirdfringediameterum, Fourthfringediamterum))
fringediameterall
summary(fringediameterall) #This shows the mean, median, minimum, and maximums through each of these data sets.
```
This can be repeated with the average diameter, however we don't need to put it in a combined group.
```{r}
summary(averagediameter)
```
We can take ANOVAs for each of these data sets as well.
```{r}
 fringediameter<- stack(fringediameterall)
fringediameter
Anova_fringediameter <- aov(values ~ ind, data = fringediameter)
summary(Anova_fringediameter)
```

