---
title: "R code"
author: "Alexis Gilliam"
date: "May 2, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Inputting the Data
To start using the program data must be present. The data being used is currently in another R Markdown document but must be added to this one as well in a readable format. There are many different approaches to this.For our purposes this data is presented as a data frame.

This first set of data show the individual fringe, diameter, wavelength, and distance.
```{r}
diffractiondata <- read.table(header=TRUE, text="
Fringe   Diameter(um) Wavelength(nm)   Distance(cm)   
1        88.64        650              2.2
1        975          650              2.0
1        43.33        650              1.5
1        58.56        640              1.0
2        86.67        650              4.5
2        86.67        650              4.5
2        52.00        650              2.5
2        58.56        640              2.0
3        82.39        650              7.1
3        90.0         650              6.5
3        51.3         650              3.8
3        79.85        640              2.2
4        84.78        650              9.2
4        91.76        650              8.5
4        54.16        650              4.8
4        83.65        640              2.8
average  79.32        648              N/A")    
```

This second data frame measures the average diameter of each hair color. 
```{r}
averagediameter <- read.table(header=TRUE, text="
Group     Average_Diameter(um)   Color
1           196.78               brown
2           96.4                 blonde
3           161.76               brown
4           85.62                brown
5           30.58                red
6           122.8                red")
```
# Plotting Data 

In the start of analysis, the data of the first data frame is analyzed in a scatter plot. The color of each of the fringes is changed to allow easy comparison.

```{r}
library(ggplot2)
ggplot(diffractiondata, aes(x = Distance.cm., y = Diameter.um.)) + 
  geom_point(aes(color = factor(Fringe)))
```

This same comparison can be drawn for the second data frame which brings into question the actually hair colors.

```{r}
library(ggplot2)
ggplot(averagediameter, aes(x = Group, y = Average_Diameter.um.)) +
  geom_point(aes(color = factor(Color)))
```

This makes it easy to compare which groups worked with which hair color and what their results are compared to each other. 